Evaluating infectious disease forecasts with allocation scoring rules

### 2023-05-23 alloscore_manuscript/allocation.tex
some progress on reviewing applications of decision theoretic forecast eval

\title{Evaluating infectious disease forecasts with allocation scoring rules}
\author{Authors}

@@

Infectious disease forecasts have been used to inform decision-making about a wide variety of measures designed to reduce disease spread and/or mitigate the severity of disease outcomes. Such decision-making applications include the allocation of limited medical supplies such as ventilators \cite{bertsimas2021predictionsCOVID}, implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials \cite{bertsimas2021predictionsCOVID}, and strategies for public health communication campaigns.

In decision-making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to. We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}. These methods have been applied with some regularity in fields such as economics and meteorology. [review previous applications of decision-theoretic evaluation to fields like economics and high level description of scoring procedures they use]

However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally. For example, Ioannidis et al. \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts. Separately, there is a thread of literature that does quantify the link between infectious disease modeling and policy making, but this work has been done outside of a forecasting context. As an example, Probert et al. \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.

In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and evaluated with standard, ``off-the-shelf'' scoring rules.
For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log-score and mean absolute errors to evaluate forecast skill \cite{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
Pandemic COVID-19 forecasts of observed cases, hospitalizations and/or deaths in the US and Europe, as reported by municipal, state, or federal surveillance systems, were evaluated using the weighted interval score (WIS, which is an approximation of the continuous ranked probability score, or CRPS), and prediction interval coverage\cite{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
Similarly, CRPS was also used to assess probabilistic forecasts of dengue incidence at the district level in Vietnam\cite{colon-gonzalez_probabilistic_2021}.
While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above had authors from public health agencies, none of them (including some for which we were co-authors) make explicit connections between forecast evaluation and how a forecast was used in practice.

@@

\bibliographystyle{unsrt}
\bibliography{allocation.bib}

>>>>

\usepackage{natbib}
\bibliographystyle{unsrtnat}

\title{Evaluating infectious disease forecasts with allocation scoring rules}
\author{Aaron Gerding, Nicholas G. Reich, Benjamin Rogers, Evan L. Ray}

@@

Infectious disease forecasts have been used to inform decision-making about a wide variety of measures designed to reduce disease spread and/or mitigate the severity of disease outcomes. Such decision-making applications include the allocation of limited medical supplies such as ventilators \citep{bertsimas2021predictionsCOVID}, implementation of social distancing measures such as stay-at-home policies, planning site selection for vaccine trials \citep{bertsimas2021predictionsCOVID}, and strategies for public health communication campaigns.

In decision-making settings where it is possible to quantify the utility or loss associated with a particular action, standard tools of decision theory provide a procedure for developing forecast scoring rules that measure the value of forecasts through the quality of the decisions that they lead to. We give an overview of these procedures in Section \ref{sec:methods.decisiontheory}. There is a large history of literature applying these ideas to obtain measures of the value of forecasts that are tied to a decision-making context, primarily in fields such as economics and finance, supply chain management, and meteorology. We review this work only briefly here, and we refer the reader to \cite{yardley2021utility_cost_forecasts} for a general overview, and to \cite{pesaran2002decision_based_eval} and \elr{TODO: identify relevant review or book style summary for meteorology} for discussions focused on applications to economics and meteorology, respectively. In finance, the value of forecasts can often be measured by the profits generated by trading decisions informed by the forecasts, perhaps adjusted for risk levels \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}. In applications to supply chain management and meteorology, the value of forecasts has typically been operationalized by considering the costs associated with decisions regarding the amount of inventory to hold or the level of protection against the impacts of extreme weather events to enact {Peter Catt (2007), Fotios Petropoulos and colleagues (2019), Nada Sanders and Gregory Graman (2009), T.N. Palmer (2002), Florian Pappenberger and colleagues (2015)}. For example, in supply chain management these decisions may incur costs related to holding inventory, labor, or providing poor service, while in meteorology we may need to balance the costs of implementing protective measures with the costs of potentially preventable weather damages. In this framework, a forecast has value if it leads to decisions with low total costs. In all of these fields, analyses have consistently found that common measures of statistical forecast accuracy do not necessarily correspond directly to measures of the value of forecasts as an input to decision-making \cite[e.g.,][]{leitch1991economicForecastEval, cenesizoglu2012returnPredictionEconValue}. % Consistently, but not uniformly

However, we are aware of only a limited body of work that explicitly attempts to measure the value of infectious disease forecasts through their impact on policy, and much of this discussion has proceeded informally. For example, \cite{ioannidis2022forecastingCOVIDfailed} discuss the possible negative consequences of inaccurate forecasts of infectious disease, but do not attempt to quantify the utility or loss incurred as a result of those forecasts. Separately, there is a thread of literature that does quantify the link between infectious disease modeling and policy making, but this work has been done outside of a forecasting context. As an example, \cite{Probert2016decisionMakingFootMouth} develop measures of the cost of actions designed to control a hypothetical outbreak of foot-and-mouth disease and use this framework to explore policy recommendations from a variety of simulation-based projection models.

In practice, probabilistic infectious disease forecasts have most often been made for observations that emerge from public health surveillance systems and evaluated with standard, ``off-the-shelf'' scoring rules.
For example, seasonal influenza forecasts in the US and dengue forecasts for Peru and Puerto Rico targeted public health surveillance measures of incidence over time and space, and used log-score and mean absolute errors to evaluate forecast skill \citep{mcgowan_collaborative_2019,reich_collaborative_2019,johansson_open_2019}.
Pandemic COVID-19 forecasts of observed cases, hospitalizations and/or deaths in the US and Europe, as reported by municipal, state, or federal surveillance systems, were evaluated using the weighted interval score (WIS, which is an approximation of the continuous ranked probability score, or CRPS), and prediction interval coverage \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.
Similarly, CRPS was also used to assess probabilistic forecasts of dengue incidence at the district level in Vietnam \citep{colon-gonzalez_probabilistic_2021}.
While some of these scores can be interpreted through the lens of decision theory, and all of the application-specific papers cited above had authors from public health agencies, none of them (including some for which we were co-authors) make explicit connections between forecast evaluation and how a forecast was used in practice.

@@

\bibliography{allocation}
